---
title: "A Questionable Interpolation"
description: |
  A few floats with interpolated positions
author:
  - name: Dewey Dunnington
    url: https://github.com/paleolimbot
date: 05-14-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
# keep warnings out of the post
library(argoFloats)
library(dplyr)
library(argodata)

temp_argo <- tempfile()
dir.create(temp_argo)
options(argoFloats.destdir = temp_argo)

knitr::opts_chunk$set(echo = TRUE)
```

These floats were [first posted](https://github.com/ArgoCanada/argoFloats/issues/224) about by [Jaimie Harbin](https://github.com/j-harbin), who noted that some interpolations are particularly questionable. This affects Canadian waters more so than others as we have a lot of ice (which prevents floats from communicating their position) and a complex coastline (which the default linear interpolation frequently crosses).

```{r}
library(argoFloats)
index1 <- subset(getIndex(), ID = '6902728')
plot(index1) 
```

```{r}
bindex <- subset(getIndex(filename = 'bgc'), ID = '6902967')
plot(bindex)
```

These points are documented as "interpolated" in the individual .nc files, identified as the `POSITION_QC` variable having a value of 8.

```{r}
library(argodata)
library(dplyr)
library(ggplot2)

profiles <- bind_rows(
  argo_global_bio_prof(),
  argo_global_prof()
) %>%
  argo_filter_float(c(6902967, 6902728)) %>%
  argo_prof_prof() %>%
  group_by(date_location, latitude, longitude) %>%
  slice(1) %>%
  ungroup()

ggplot(profiles, aes(longitude, latitude)) +
  geom_point(aes(col = position_qc)) +
  theme_void()
```

More detail about the profile positions:

```{r, layout="l-body-outset"}
rmarkdown::paged_table(profiles)
```


```{r, include = FALSE}
# clean up topo files so they don't end up on GitHub
unlink(list.files(".", "^topo", full.names = T))
```
