---
title: "Canadian Argo: Future Planning and Coordination"
description: |
  Contains: info on deployments, planning, float inventory, and expected orders.
author:
  - name: Christopher Gordon
    url: https://github.com/cgrdn
date: 2024-10-31
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(reticulate)
use_condaenv("py39")
```

```{python imports and load dataframe}

import numpy as np
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt

df = pd.read_csv('../../deployment/ArgoCanada.csv')
df = df.loc[df.Scientist != 'Marcel Babin']
df['Date'] = [
  pd.Timestamp(
    year=int(row.Datestring.split('/')[0]),
    month=int(row.Datestring.split('/')[1]),
    day=int(row.Datestring.split('/')[2]),
    hour=int(row.Time.split(':')[0]),
    minute=int(row.Time.split(':')[1])
  ) if not pd.isna(row.Datestring) else pd.NaT for _, row in df.iterrows() 
]
df['Year'] = [d.year for d in df.Date]

n_deployed = df.loc[df.Date.notna()].shape[0]
earliest = df.Date.min().year
n2022 = df.loc[df.Year == 2022].shape[0]
n2023 = df.loc[df.Year == 2023].shape[0]
n2024 = df.loc[df.Year == 2024].shape[0]

deployed = df.loc[df.Date.notna()]

stock = df.loc[df.Date.isna()]
stock = stock.reset_index().loc[:, ['WMO', 'Serial', 'Model']]
nstock = stock.shape[0]

order = pd.DataFrame(
  dict(
    Institute=['DFO', 'DFO', 'Dal', 'Dal'],
    Number=[7, 7, 10, 17],
    Float=['APEX', 'APEX', 'PROVOR CTS5', 'PROVOR CTS5'],
    Sensors=['SBE CTD', 'RBR CTD', 'SBE CTD, DO Optope, RBR Tridente, OCR', 'SBE CTD, DO Optope, RBR Tridente, OCR, Trios OPUS']
  )
)

```

Since `r py$earliest`, Argo Canada has deployed `r py$n_deployed` floats. In recent years, Canada has deployed `r py$n2022` in 2022, `r py$n2023` in 2023, and `r py$n2024` so far in 2024, with 2 planned deployments remaining for the year. 

```{python deployments graph, layout="l-body"}

deployments = pd.DataFrame(
  dict(
    Year = deployed.Year.astype(int).unique(),
    Deployments = [deployed.loc[df.Year == y].shape[0] for y in deployed.Year.unique()],
    Cumulative = [deployed.loc[df.Year <= y].shape[0] for y in deployed.Year.unique()]
  )
)

g = sns.histplot(deployed, x='Year', hue='Model', multiple='stack', bins=np.arange(2011.5, 2025.5, 1), palette='colorblind')
ax = g.axes.twinx()
sns.lineplot(deployments, x='Year', y='Cumulative', ax=ax)
plt.show()

```

From the rightmost bar it is clear that we are now deploying a diverse set of floats, with additional models/sensors to be added in the next few years.

The following `r py$nstock` floats are left "in stock", including P43205-22CA004 and P53875-23CA001 which are planned for deployment later this year:

```{r table 1, layout="l-body shaded"}
knitr::kable(py$stock)
```

Finally, the following floats are on order:

```{r table 2, layout="l-body shaded"}
knitr::kable(py$order)
```
