---
title: "Deployment Planning"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(reticulate)
library(leaflet)
# use_condaenv("py39")
```

This page is used for Canadian Argo groups to coordinate deployment planning, in addition to [OceanOps](https://www.ocean-ops.org/board?t=argo). To submit plans to the website, see instructions on the [github page](https://github.com/ArgoCanada/blog/tree/master/deployment).

The map below shows completed and planned deployments for 2023, coloured by institution. Dots with a buoy symbol are locations of completed deployments (deployment location, not most recent profile location), while dots with the loading symbol are planned deployments.

```{python data}
import pandas as pd

df = pd.read_csv('deployment/canada_deployments.csv')
df['DEPLOYMENT DATE'] = df['DEPLOYMENT DATE'].apply(pd.Timestamp)
df = df.loc[df['DEPLOYMENT DATE'] > pd.Timestamp('01-01-2023')]
df = df.sort_values('DEPLOYMENT DATE')
df['DEPLOYMENT DATE'] = [d.strftime('%d %b, %Y') for d in df['DEPLOYMENT DATE']]
df['WMO'] = [str(int(w)) if pd.notna(w) else '' for w in df.WMO]
df['IMEI'] = [str(int(i)) if pd.notna(i) else '' for i in df.IMEI]
df['SERIAL NUMBER'] = [s if pd.notna(s) else '' for s in df['SERIAL NUMBER']]

mapper = {
  'PROGRAM':'Program',
  'INSTITUTE':'Institute',
  'STATUS':'Status',
  'MODEL':'Model',
  'DEPLOYMENT DATE':'Date',
  'DEPLOYMENT LAT':'Latitude',
  'DEPLOYMENT LON':'Longitude',
  'DEPLOYMENT SHIP':'Ship',
  'IMEI':'IMEI',
  'REF':'WMO',
  'SERIAL NUMBER':'Serial No.',
}

df = df.rename(columns=mapper)
df = df[mapper.values]
df = df.reset_index().drop('index', axis=1)

recent = df.loc[df.Status == 'OPERATIONAL'].drop('Status', axis=1).reset_index().drop('index', axis=1)
plan   = df.loc[df.Status != 'OPERATIONAL'].drop('Status', axis=1).reset_index().drop('index', axis=1)
```

```{r map, layout="l-page-outset"}
pal <- colorFactor(
  c("green", "purple", "blue", "yellow"),
  levels = c("BIO", "IOS", "ONC", "Dal")
)
icons <- iconList(
  OPERATIONAL = makeIcon("images/buoy.png", "buoy.png", 16, 16),
  CONFIRMED = makeIcon("images/loading.png", "loading.png", 16, 16)
)

leaflet() %>%
  addTiles() %>%
    setView(lng = -60, lat = 45, zoom = 2) %>%
    addCircleMarkers(
      data = py$df,
      lat = py$df$Latitude,
      lng = py$df$Longitude,
      radius = 12,
      stroke = FALSE,
      fillColor = ~pal(py$df$Institute),
      fillOpacity = 0.7,
      popup = paste0(
        "WMO: ", py$df$WMO,
        "<br/>",
        "Model: ", py$df$Model,
        "<br/>",
        "Program: ", py$df$Program, ", ", py$df$Institute,
        "<br/>",
        "Deployment Date: ", py$df$Date,
        "<br/>",
        "Ship: ", py$df$Ship
      )
    ) %>%
    addMarkers(
      data = py$df,
      lat = py$df$Latitude,
      lng = py$df$Longitude,
      icon = ~icons[py$df$Status],
      popup = paste0(
        "WMO: ", py$df$WMO,
        "<br/>",
        "Model: ", py$df$Model,
        "<br/>",
        "Program: ", py$df$Program, ", ", py$df$Institute,
        "<br/>",
        "Deployment Date: ", py$df$Date,
        "<br/>",
        "Ship: ", py$df$Ship
      )
    )
```

## Deployments since Jan 1, 2023

```{r table_1, layout="l-page-outset shaded"}
knitr::kable(py$recent)
```
## Planned Deployments

```{r table_2, layout="l-page-outset shaded"}
knitr::kable(py$plan)
```

### Other Deployment Information

Although there is not yet a designated cruise or location, Argo Canada is sending 6 ARVOR floats to Cape Town, South Africa via the RRS Discovery. These floats will be deployed in the Western Indian Ocean where there is a significant gap in the Argo array.

## Inventory

```{python inventory}
stock = pd.read_csv('deployment/inventory.csv')

remain = pd.DataFrame({
    'Program':stock.Program,
    'Institute':stock.Institute,
    'ARVOR':stock['ARVOR'] + stock['ARVOR+DO'] + stock['ARVOR-RBR'],
    'Deep ARVOR':stock['Deep ARVOR'],
    'PROVOR':stock['PROVOR']
})

remain = remain.set_index(['Program', 'Institute'], drop=False)

for i in plan.index:
    ix = (plan.loc[i].Program, plan.loc[i].Institute)
    f = plan.loc[i].Model
    remain.loc[ix, f] -= 1

remain.loc[('Argo Canada','BIO'), 'ARVOR'] -= 6 # account for Cape Town
```

The table below shows the number of "in stock" floats for each program/institution.

```{r table_3, layout="l-body-outset shaded"}
knitr::kable(py$stock)
```

This table shows what the standing stock of floats will be for each program/institution will be following the completion of the deployments listed in the "Planned Deployments" section. For BIO, the 6 floats being sent to Cape Town are also being accounted for in this table. For simplicity, all types of ARVOR models (except deep) are combined here.

```{r table_4, layout="l-body-outset shaded"}
knitr::kable(py$remain)
```
