---
title: "Argo Canada Data Management Team Updates - June 21, 2024"
description: |
  Summary of updates for meeting taking place on June 21, 2024.
author:
  - name: Christopher Gordon
    url: https://github.com/cgrdn
date: 2024-06-20
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(leaflet)
library(emoji)
library(reticulate)
```

```{python imports and setup, include=FALSE}
### imports and setup

import argopy
import argopandas as argo

from pathlib import Path
from netCDF4 import Dataset
import numpy as np

import matplotlib.pyplot as plt
import matplotlib.path as mpath
import cmocean.cm as cmo

import pandas as pd
import seaborn as sns
sns.set(style="ticks", palette="colorblind")
```

## Recent Deployments

Recent deployments (since May 1) include 5 floats in the Lab Sea (3 Dal CTS5, 2 BIO CTS4), 1 float deployed from the sailing vessel Oceanolog (read more [here](https://sailforscience.com/a-day/)), and 2 floats deployed along line P (1 deep ARVOR, 1 CTS4 w/ Nitrate).

Going a little further back, in April 4 floats were deployed, 2 on the AZMP (BIO CTS4) and 2 southeast of South Africa (ARVOR).

Dal floats: profiled daily while ship was in the area, now on 5-day cycles. Parking at 2000m to use as a pseudo-sediment trap via the BBP parking hook (see [here](https://argocanada.github.io/blog/posts/2023-12-05-implementing-updated-backscatter-rtqc/#parking-hook-test) for an example)

Nitrate float (4902679) is not on GDAC yet (Nitrate processing in progress), profiling daily?

```{python data, include=FALSE}
import pandas as pd

df = pd.read_csv('../../deployment/canada_deployments.csv')
df['DEPLOYMENT DATE'] = df['DEPLOYMENT DATE'].apply(pd.Timestamp)
df = df.loc[df['DEPLOYMENT DATE'] > pd.Timestamp('05-01-2024')]
df = df.sort_values('DEPLOYMENT DATE')
df['DEPLOYMENT DATE'] = [d.strftime('%d %b, %Y') for d in df['DEPLOYMENT DATE']]
df['REF'] = [str(int(w)) if pd.notna(w) else '' for w in df.REF]
df['IMEI'] = [str(int(i)) if pd.notna(i) else '' for i in df.IMEI]
df['SERIAL NUMBER'] = [s if pd.notna(s) else '' for s in df['SERIAL NUMBER']]
df['DEPLOYMENT SHIP'] = [s if pd.notna(s) else '' for s in df['DEPLOYMENT SHIP']]

mapper = {
  'PROGRAM':'Program',
  'INSTITUTE':'Institute',
  'STATUS':'Status',
  'MODEL':'Model',
  'MODEL_DETAIL':'MODEL_DETAIL',
  'DEPLOYMENT DATE':'Date',
  'DEPLOYMENT LAT':'Latitude',
  'DEPLOYMENT LON':'Longitude',
  'DEPLOYMENT SHIP':'Ship',
  'IMEI':'IMEI',
  'REF':'WMO',
  'SERIAL NUMBER':'Serial No.',
}

df = df.rename(columns=mapper)
df = df[mapper.values()]
df = df.reset_index().drop('index', axis=1)

mdf = pd.read_csv("../../deployment/ship_info.csv")

recent = df.loc[df.Status == 'OPERATIONAL'].drop('Status', axis=1).reset_index().drop('index', axis=1)
recent_table = recent.drop('MODEL_DETAIL', axis=1)
plan   = df.loc[df.Status == 'CONFIRMED'].drop('Status', axis=1).reset_index().drop('index', axis=1)
plan_table = plan.drop('MODEL_DETAIL', axis=1)
```

```{r map, layout="l-page-outset", echo=FALSE}
pal <- colorFactor(
  c("green", "purple", "blue", "yellow","green"),
  levels = c("BIO", "IOS", "ONC", "Dal","SAEON")
)
icons <- iconList(
  OPERATIONAL = makeIcon("../../images/buoy.png", "../../images/buoy.png", 16, 16),
  CONFIRMED = makeIcon("../../images/loading.png", "../../images/loading.png", 16, 16),
  FAILED = makeIcon("../../images/x.png", "../../images/x.png", 16, 16)
)

leaflet() %>%
  addTiles() %>%
    setView(lng = -60, lat = 45, zoom = 2) %>%
    addCircleMarkers(
      data = py$df,
      lat = py$df$Latitude,
      lng = py$df$Longitude,
      radius = 12,
      stroke = FALSE,
      fillColor = ~pal(py$df$Institute),
      fillOpacity = 0.7,
      popup = paste0(
        "WMO: ", py$df$WMO,
        "<br/>",
        "Model: ", py$df$Model,
        "<br/>",
        "Program: ", py$df$Program, ", ", py$df$Institute,
        "<br/>",
        "Deployment Date: ", py$df$Date,
        "<br/>",
        "Ship: ", py$df$Ship
      )
    ) %>%
    addMarkers(
      data = py$df,
      lat = py$df$Latitude,
      lng = py$df$Longitude,
      icon = ~icons[py$df$Status],
      popup = paste0(
        "WMO: ", py$df$WMO,
        "<br/>",
        "Model: ", py$df$Model,
        "<br/>",
        "Program: ", py$df$Program, ", ", py$df$Institute,
        "<br/>",
        "Deployment Date: ", py$df$Date,
        "<br/>",
        "Ship: ", py$df$Ship
      )
    )

knitr::kable(py$recent_table)
```

## Arctic & Antarctic Floats

```{python arctic summary, include=FALSE}
# wmo numbers of the floats
beaufort_wmos = [4902610, 4902611]
ross_wmos = [4902664, 4902665, 4902667, 4902668, 4902669]

# grab Argo index for each group
index = argopy.ArgoIndex().load()
beaufort_ix = index.search_wmo(beaufort_wmos).to_dataframe()
ross_ix = index.search_wmo(ross_wmos).to_dataframe()

hl = index.search_wmo(beaufort_wmos + ross_wmos).to_dataframe()

table = pd.DataFrame({
  'WMO':[wmo for wmo in hl.wmo.unique()],
  'Basin':['Beaufort Sea' if wmo in beaufort_ix.wmo.values else 'Ross Sea' for wmo in hl.wmo.unique()],
  'Deployment Date':[hl.loc[hl.wmo == wmo].date.min().strftime('%b %d, %Y') for wmo in hl.wmo.unique()],
  'Last Cycle':[hl.loc[hl.wmo == wmo].cyc.max() for wmo in hl.wmo.unique()],
  'Last Cycle Date':[hl.loc[hl.wmo == wmo].date.max().strftime('%b %d, %Y') for wmo in hl.wmo.unique()],
})
  
```

```{r table, include=TRUE, echo=FALSE, layout="l-body-outset shaded"}
knitr::kable(py$table)
```

Beaufort:

- Both Beaufort Sea floats have resurfaced! 4902610 not on GDAC yet.
- ISA does not seem to have prevented floats from approaching/bumping into the ice
- Lost some data - ex. 4902611 did 3 cycles before going under ice, but newly sent data starts at cycle 13: reduce vertical resolution before next season?

Ross:

- Most floats have not reported since near deployment time, but 4902664 just reported on June 16, not on GDAC yet


```{python beaufort float data, include=FALSE, echo=FALSE}

ix = argo.float(4902611).prof
data = ix.levels[['PRES', 'TEMP', 'PSAL']]

fig, axes = plt.subplots(1, 2, sharey=True)
for ax, v in zip(axes, ['TEMP', 'PSAL']):
  sns.lineplot(
    data=data, x=v, y='PRES', hue='file', style='N_PROF',
    legend=False, sort=False, estimator=None,
    palette='Blues', ax=ax
  )

ax.set_ylim((2050, -50))
```

```{python plot show}
plt.show()
```

## Nitrate

Writing of Nitrate RTQC code is in progress. Official documentation is [here](https://archimer.ifremer.fr/doc/00732/84370/89380.pdf). RTQC for Nitrate includes: 

- global range test
- spike test
- stuck value test
- grey list test
- nitrate specific tests

Specific tests include checking for saturation of the spectrophotometer, absorbance at 240nm (evaluate sensor optics) and error fitting to a model. There is also a check for high bisulfide concentrations such as in the black sea, but these are special cases and likely only going to be evaluated by expert researchers. 

Like oxygen, unadjusted nitrate is flagged as 3 by default. If TEMP_QC is 4, NITRATE_QC is also 4, and if PSAL_QC is 4, NITRATE_QC should be 3 as well, including for adjusted, but this can be reviewed at the DMQC step. 

## Near Surface Sampling

At the most recent AST Brian King presented on an evaluation of Argo floats sampling all the way to the surface. The defauly configuration of the CTD pump for NKE floats is to turn off the pump at 5dbar. I am in the process of changing this parameter to 2dbar for all floats and will use as the default going forward. 
