---
title: "Labrador Shelf DFO Argo Float"
description: |
  A quick overview of a DFO Argo float that survived a winter drifting under the ice on the Labrador Shelf
author:
  - name: Clark Richards
date: 05-14-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = "png", dev.args = list(type = "cairo-png"))
```

## Introduction

In July 2017, A DFO Argo float (of NOVA type, containing a pumped SBE CTD) was deployed on the Labrador Shelf, as part of an exercises related to the DFO/ECCC/DND CONCEPTS program (Canadian Operational Network of Coupled Environmental PredicTion Systems). The float was programmed to execute daily profiles between the surface and the bottom until it exhausted its batteries.

The float, with WMO ID 4902426, executed 389 total profiles, surviving a winter season under the ice. Float data and trajectory can be explored through the EuroArgo "Fleet monitoring" site, found at: [https://fleetmonitoring.euro-argo.eu/float/4902426](https://fleetmonitoring.euro-argo.eu/float/4902426), or by downloading the netCDF files directly from [https://data-argo.ifremer.fr/dac/meds/4902426/](https://data-argo.ifremer.fr/dac/meds/4902426/)

## Data summary

A summary of the data found in the combined profile file `4902426_prof.nc` is provided below (using the [R oce](https://dankelley.github.io/oce/) package developed by myself and Dan Kelley). Another approach, which permits individual access to the single-profile files provided on the Argo DAC, is to use the *in development* (but quite mature) [`argoFloats`](https://argocanada.github.io/argoFloats/index.html) package.

To see a crude map of the float profile locations:
```{r}
library(oce)
a <- read.oce('4902426_prof.nc')
plot(a)
```

For a quick "section"-style plot of the temperature, salinity, and density data:
```{r}
s <- as.section(handleFlags(a)) # convert the float object to a section object,
                                # applying the argo QC flags
library(cmocean) # for nice colormaps
plot(s, which='temperature', xtype='time', ztype='image', col=cmocean('thermal'))
plot(s, which='salinity', xtype='time', ztype='image', col=cmocean('haline'))
plot(s, which='sigmaTheta', xtype='time', ztype='image', col=cmocean('dense'))
```

