---
title: "2024"
description: |
  Argo Canada Deployment Summary for 2024
author:
  - name: Christopher Gordon
    url: https://github.com/cgrdn
date: 2025-01-14
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(reticulate)
library(leaflet)
# use_condaenv("py39")
```

The map below shows completed deployments for 2023, coloured by institution. Locations represent deployment location, not last profile. Click the dots for information about the float.

```{python data}
import pandas as pd

df = pd.read_csv('../../deployment/canada_deployments.csv')
df['DEPLOYMENT DATE'] = df['DEPLOYMENT DATE'].apply(pd.Timestamp)
df = df.loc[df['DEPLOYMENT DATE'] > pd.Timestamp('31-12-2023')]
df = df.loc[df['DEPLOYMENT DATE'] < pd.Timestamp('01-01-2025')]
df = df.sort_values('DEPLOYMENT DATE')
df['DEPLOYMENT DATE'] = [d.strftime('%d %b, %Y') for d in df['DEPLOYMENT DATE']]
df['REF'] = [str(int(w)) if pd.notna(w) else '' for w in df.REF]
df['IMEI'] = [str(int(i)) if pd.notna(i) else '' for i in df.IMEI]
df['SERIAL NUMBER'] = [s if pd.notna(s) else '' for s in df['SERIAL NUMBER']]

mapper = {
  'PROGRAM':'Program',
  'INSTITUTE':'Institute',
  'STATUS':'Status',
  'MODEL':'Model',
  'MODEL_DETAIL':'Detail',
  'BASIN':'Basin',
  'DEPLOYMENT DATE':'Date',
  'DEPLOYMENT LAT':'Latitude',
  'DEPLOYMENT LON':'Longitude',
  'DEPLOYMENT SHIP':'Ship',
  'IMEI':'IMEI',
  'REF':'WMO',
  'SERIAL NUMBER':'Serial No.',
}

df = df.rename(columns=mapper)
df = df[mapper.values()]
df = df.reset_index().drop('index', axis=1)
table = df.drop(['Program', 'Detail', 'IMEI', ], axis=1)

mdf = pd.read_csv("../../deployment/ship_info.csv")

n_floats = df.shape[0]
n_sbe  = df.loc[df.Detail == 'ARVOR_SBE'].shape[0]
n_rbr  = df.loc[df.Detail == 'ARVOR_RBR'].shape[0]
n_do   = df.loc[df.Detail == 'ARVOR+DO'].shape[0]
n_deep = df.loc[df.Detail == 'ARVOR_D'].shape[0]
n_cts4 = df.loc[df.Detail == 'PROVOR_CTS4'].shape[0]+df.loc[df.Detail == 'PROVOR_CTS4+SUNA'].shape[0]+df.loc[df.Detail == 'PROVOR_CTS4+pH'].shape[0]
n_cts5 = df.loc[df.Detail == 'PROVOR_CTS5'].shape[0]

n_atl  = df.loc[df.Basin == 'ATL'].shape[0]
n_pac  = df.loc[df.Basin == 'PAC'].shape[0]
n_arc  = df.loc[df.Basin == 'ARC'].shape[0]
n_sou  = df.loc[df.Basin == 'SOU'].shape[0]
n_ind  = df.loc[df.Basin == 'IND'].shape[0]
```

```{r map, layout="l-page-outset"}
pal <- colorFactor(
  c("green", "purple", "blue", "yellow", "green"),
  levels = c("BIO", "IOS", "ONC", "Dal", "SAEON")
)
icons <- iconList(
  OPERATIONAL = makeIcon("../../images/buoy.png", "../../images/buoy.png", 16, 16),
  CONFIRMED = makeIcon("../../images/loading.png", "../../images/loading.png", 16, 16),
  FAILED = makeIcon("../../images/x.png", "../../images/x.png", 16, 16)
)

leaflet() %>%
  addTiles() %>%
    setView(lng = -60, lat = 45, zoom = 2) %>%
    addCircleMarkers(
      data = py$df,
      lat = py$df$Latitude,
      lng = py$df$Longitude,
      radius = 12,
      stroke = FALSE,
      fillColor = ~pal(py$df$Institute),
      fillOpacity = 0.7,
      popup = paste0(
        "WMO: ", py$df$WMO,
        "<br/>",
        "Model: ", py$df$Model,
        "<br/>",
        "Program: ", py$df$Program, ", ", py$df$Institute,
        "<br/>",
        "Deployment Date: ", py$df$Date,
        "<br/>",
        "Ship: ", py$df$Ship
      )
    ) %>%
    addMarkers(
      data = py$df,
      lat = py$df$Latitude,
      lng = py$df$Longitude,
      icon = ~icons[py$df$Status],
      popup = paste0(
        "WMO: ", py$df$WMO,
        "<br/>",
        "Model: ", py$df$Model,
        "<br/>",
        "Program: ", py$df$Program, ", ", py$df$Institute,
        "<br/>",
        "Deployment Date: ", py$df$Date,
        "<br/>",
        "Ship: ", py$df$Ship
      )
    )
```

## Deployment Summary

In total, `r py$n_floats` floats were deployed in 2024. `r py$n_rbr + py$n_sbe` core ARVOR floats (`r py$n_sbe` SBE, `r py$n_rbr` RBR), `r py$n_do` Oxygen-only ARVOR floats, `r py$n_deep` deep-ARVOR floats with DOXY sensors, and `r py$n_cts4 + py$n_cts5` PROVOR BGC floats. Of the BGC floats, 8 were 3-sensor floats with DOXY, CHLA, and BBP, 3 were 4-sensor floats with DOXY, CHLA, BBP, and radiometry, and 2 were 4-sensor floats with DOXY, CHLA, BBP, and NITRATE. 

By ocean basin, `r py$n_atl` floats were deployed in the Atlantic, `r py$n_pac`  in the Pacific, `r py$n_arc`  in the Arctic, `r py$n_ind` in the Indian, and `r py$n_sou` in the Southern Ocean.

A complete list of deployed floats can be found below. 

```{r table_1, layout="l-page-outset shaded"}
assign_country <- function(name) {
  cty <- py$mdf$Country[py$mdf$Ship == name]
  paste(emoji::emoji_name[paste0("flag_", cty)], name)
}

link_efm <- function(wmo) {
  paste0("[",wmo,"](https://fleetmonitoring.euro-argo.eu/float/",wmo,")")
}

py$table$Ship <- lapply(py$table$Ship, assign_country)
py$table$WMO <- lapply(py$table$WMO, link_efm)
knitr::kable(py$table)
```

## Notes from the year

- One float ([4902609](https://fleetmonitoring.euro-argo.eu/float/4902609)) was deployed from the sailing vessel Oceanolog as part of Sail for Science - read about it here: [Sail for Science: A-day](https://sailforscience.com/a-day/)
- One deployed float (4902565) in the Beaufort Sea never properly activated and/or returned any technical or profile data. 
- Multiple floats in the Ross Sea, Beaufort Sea, and even one in Baffin Bay ([4902531](https://fleetmonitoring.euro-argo.eu/float/4902531), deployed in 2021!) emerged from under ice to report profiles.
- One float ARVOR-SBE experienced vacuum pressure issues and was not deployed, returned to NKE, and repaired. Currently this float is being shipped back to IOS. 
- NKE & RBR recalled 2 ARVOR-RBR floats for investigation as a number of this model across the Argo network were experiencing a rapid power drawdown resulting in premature end of float life. No systematic problem was discovered and these floats are being returned to IOS, ready to deploy. 
- One float ([4902653](https://fleetmonitoring.euro-argo.eu/float/4902653)) deployed off South Africa failed after 4 profiles, reason still to be determined.
