---
title: "Checking Argo Files"
description: |
  How to run the Argo file checker on any platform
author:
  - name: Dewey Dunnington
    url: https://github.com/paleolimbot
date: 06-23-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In preparation for writing some real-time quality control code, the question arose of how to check modified Argo files to make sure that they conform to the specification in the [Argo User's Manual](https://doi.org/10.13155/29825). The [official tool is hosted on Ifremer](https://doi.org/10.17882/45538) and is written in Java. When the latest version is downloaded you can run a shell command that will check one or more files. Running a `bash` shell, this can be done in a few lines:

``` bash
# download and unpack the tool in the current working directory
curl https://www.seanoe.org/data/00344/45538/data/83774.tar.gz | tar -xz

# set the working directory
cd format_control_1-17

# download a test Argo file to check
curl -o R4902533_001.nc \
    https://data-argo.ifremer.fr/dac/meds/4902533/profiles/R4902533_001.nc

# ...and check it
java -cp ./resources:./jar/formatcheckerClassic-1.17-jar-with-dependencies.jar \
    -Dapplication.properties=application.properties \
    -Dfile.encoding=UTF8 \
    oco.FormatControl \
    R4902533_001.nc
```

    <?xml version="1.0"?>
    <coriolis_function_report>
            <function>CO-03-08-03</function>
            <comment>Control file data format</comment>
            <date>23/06/2021 12:50:59</date>
            <application_version>1.17</application_version>
            <netcdf_file>R4902533_001.nc</netcdf_file>
            <rules_file>Argo_Prof_c_v3.1_AUM_3.1_20201104.xml</rules_file>
            <title>Argo float vertical profile</title>
            <user_manual_version>3.1</user_manual_version>
            <data_type>Argo profile</data_type>
            <format_version>3.1</format_version>
            <file_error>The variable "LATITUDE" is not correct: attribute "reference" forbidden</file_error>
            <file_error>The variable "LATITUDE" is not correct: attribute "coordinate_reference_frame" forbidden</file_error>
            <file_error>The variable "LONGITUDE" is not correct: attribute "reference" forbidden</file_error>
            <file_error>The variable "LONGITUDE" is not correct: attribute "coordinate_reference_frame" forbidden</file_error>
            <file_error>The optional variable "PRES" is not correct: attribut "coordinate_reference_frame" forbidden</file_error>
            <file_error>The value of the attribute of variable "PRES_ADJUSTED:axis" is not correct: "Z" expected</file_error>
            <file_compliant>no</file_compliant>
            <status>ok</status>
    </coriolis_function_report>


If you're on Windows and you're running in PowerShell or good ol' `cmd.exe`, you can just run `command.bat  R4902533_001.nc` (although you'll need to download and extract the tool separately); a shell wrapper for Linux is also distributed but appears to hard-code the location of `java` to the developer's computer so you won't be able to run it without some modification. If you're on Windows and running Git Bash, you'll need to replace the `:` separating the class paths with `\;` because that's how the Java interpreter expects paths to be separated on Windows (and because `;` is a special character in `bash` so you need to escape it with `\`).

If you're writing QC code in Python you can run the tool directly using the `subprocess` module from the standard library.

``` python
import subprocess
import os
import tempfile

# handle the platform-dependence of the class-path separator
if os.name == 'nt':
    classpath_sep = ';'
else:
    classpath_sep = ':'

class_path_rel = ('./resources', './jar/formatcheckerClassic-1.17-jar-with-dependencies.jar')
class_path = classpath_sep.join(class_path_rel)

# construct arguments as a list()
args = [
    'java', '-cp', class_path,
    '-Dapplication.properties=application.properties',
    '-Dfile.encoding=UTF8',
    'oco.FormatControl',
    'R4902533_001.nc'
]

result = subprocess.run(args, cwd='format_control_1-17', capture_output=True)
result.check_returncode()
```

This will run the tool and check for a non-zero status code (e.g., `java` fails to start). The bytes of the xml are available as `result.stdout` (if you want to output directly to stdout you can do so by omitting `capture_output=True`). You can then parse the results using the `xml.etree.ElementTree` class:

``` python
from xml.etree import ElementTree
import io

root = ElementTree.parse(io.BytesIO(result.stdout)).getroot()
errors = [el.text for el in root.findall('file_error')]
errors
```

    ['The variable "LATITUDE" is not correct: attribute "reference" forbidden',
     'The variable "LATITUDE" is not correct: attribute "coordinate_reference_frame" forbidden',
     'The variable "LONGITUDE" is not correct: attribute "reference" forbidden',
     'The variable "LONGITUDE" is not correct: attribute "coordinate_reference_frame" forbidden',
     'The optional variable "PRES" is not correct: attribut "coordinate_reference_frame" forbidden',
     'The value of the attribute of variable "PRES_ADJUSTED:axis" is not correct: "Z" expected']

You can use a similar trick to run and parse the results in R with the help of the [processx](https://processx.r-lib.org/) and [xml2](https://xml2.r-lib.org/) packages.

```{r}
classpath_sep <- if (Sys.info()["sysname"] == "Windows") ";" else ":"
classpath <- paste(
  "./resources",
  "./jar/formatcheckerClassic-1.17-jar-with-dependencies.jar",
  sep = classpath_sep
)

args <- c(
  "-cp", classpath,
  "-Dapplication.properties=application.properties",
  "-Dfile.encoding=UTF8",
  "oco.FormatControl",
  "R4902533_001.nc"
)
result <- processx::run("java", args, wd = "format_control_1-17")

root <- xml2::read_xml(result$stdout)
errors <- xml2::xml_text(xml2::xml_find_all(root, "file_error"))
errors
```

There are some complexities that aren't handled by the simple cases above. The tool and/or the rules used to determine what constitutes a `<file_error>` are updated several times a year and keeping the tool up-to-date requires a manual check if it's installed above. These bits of code also assume that when you type `java` at a terminal that you actually get a Java interpreter! This is not always the case and configuring a Java VM can be complex.

To solve these issues I put together a proof-of-concept [Python tool](https://github.com/paleolimbot/docker-images/blob/main/argo-checker/argo-checker.py) + [Docker image](https://github.com/paleolimbot/docker-images/tree/main/argo-checker#argo-checker) that boil the above steps down to a one-liner:

``` bash
# once per computer: docker pull paleolimbot/argo-checker
curl -s https://data-argo.ifremer.fr/dac/meds/4902533/profiles/R4902533_001.nc | \
    docker run --rm paleolimbot/argo-checker --update check > result.xml
```

    Searching for installed tool in '/argo-checker'
    Installed tool found at '/argo-checker/tool_83774'.
    Checking for newer version, as requested...
    Checking for latest tool at <https://doi.org/10.17882/45538>...
    Latest tool source is located at <https://www.seanoe.org/data/00344/45538/data/83774.tar.gz>.
    Version is latest version.
    Running 'java -cp ./resources:./jar/formatcheckerClassic-1.17-jar-with-dependencies.jar -Dapplication.properties=application.properties -Dfile.encoding=UTF8 oco.FormatControl /tmp/tmpytrg3qyn.nc'

You can then process 'result.xml' using whatever tool you'd like! Probably a more robust option would be to rewrite the Java tool in Python, but that's a battle for another day.
